The Unicode Cookbook \linebreak for Linguists

### 2018-06-05 book/chapters/introduction.tex
fix en dashes

From those bisignal encodings, true \textsc{binary codes} developed with a fixed
length of signals per character. In such systems only a single separator between
signals is needed, because the separation between characters can be established
by counting until a fixed number of signals has passed.\footnote{Of course, no
explicit separator is needed at all when the timing of the signals is known, which is
the principle used in all modern telecommunication systems. An important modern
consideration is also how to know where to start counting when you did not catch
the start of a message, something that is known in Unicode as \textsc{self
synchronization}.} In the context of electric telegraphy, such a binary code
system was first established by Émile Baudot in 1870, using a fixed combination
of five signals for each written character.\footnote{True binary codes have a
longer history, going at least back to the Baconian cipher devised by Francis
Bacon in 1605. However, the proposal by Baudot was the quintessential proposal
leading to all modern systems.} There are $2^5 = 32$ possible combination when
using five binary signals; an encoding today designated as 5-bit. These
codes are sufficient for all Latin letters, but of course they do not suffice
for all written symbols, including punctuation and digits. As a solution, the
Baudot code uses a so-called shift character, which signifies that from
that point onwards---until shifted back---a different encoding is used, allowing
for yet another set of 32 codes. In effect, this means that the Baudot code, and
the \textsc{International Telegraph Alphabet} (ITA) derived from it, had an
extra bit of information, so the encoding is actually 6-bit (with $2^6
= 64$ different possible characters). For decades, this encoding was the
standard for all telegraphy and it is still in limited use today.

>>>>

From those bisignal encodings, true \textsc{binary codes} developed with a fixed
length of signals per character. In such systems only a single separator between
signals is needed, because the separation between characters can be established
by counting until a fixed number of signals has passed.\footnote{Of course, no
explicit separator is needed at all when the timing of the signals is known, which is
the principle used in all modern telecommunication systems. An important modern
consideration is also how to know where to start counting when you did not catch
the start of a message, something that is known in Unicode as \textsc{self
synchronization}.} In the context of electric telegraphy, such a binary code
system was first established by Émile Baudot in 1870, using a fixed combination
of five signals for each written character.\footnote{True binary codes have a
longer history, going at least back to the Baconian cipher devised by Francis
Bacon in 1605. However, the proposal by Baudot was the quintessential proposal
leading to all modern systems.} There are $2^5 = 32$ possible combination when
using five binary signals; an encoding today designated as 5-bit. These
codes are sufficient for all Latin letters, but of course they do not suffice
for all written symbols, including punctuation and digits. As a solution, the
Baudot code uses a so-called shift character, which signifies that from
that point onwards -- until shifted back -- a different encoding is used, allowing
for yet another set of 32 codes. In effect, this means that the Baudot code, and
the \textsc{International Telegraph Alphabet} (ITA) derived from it, had an
extra bit of information, so the encoding is actually 6-bit (with $2^6
= 64$ different possible characters). For decades, this encoding was the
standard for all telegraphy and it is still in limited use today.
